Dynamic Adaptive Saturated Sliding Mode Control For Deployment of Tethered Satellite System

### 2016-10-26 manuscript.tex
20161026

\bm\tau = -\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1})),\label{eq:slider control}

@@

\frac{d}{dt}F(t) = \delta\varphi S_{r2}^T K_1K_de^{-2K_1^{-1}F^{-1}(t)}\bm{sign}(S_{r2})

@@

\textcolor{red}{$\varphi$ is the boundedness of the lumped uncertainty. $\gamma$ is a small or  time-varying positive scalar, $\varphi>1$ and $K_d$ is the designer parameter, such that the deployment dynamics can be regulated as desired.}

@@

It is obvious that $S_{q_1}$ is bounded, or exactly saturated, hence $S_{r_1}$ is also bounded if the dynamics can be controlled as desired. In addition, one can select a suitable positive scalar $\gamma$ such that $\gamma\int^t_{t_0}S_{q1}(\varsigma) d\varsigma\in[-1,1]$.\par

@@

H(q)\dot S_{r2} = -\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))-C(q,\dot q)S_{r2}-Y_{r2}\Theta.

@@

&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TC(q,\dot q)S_{r2}+\frac{1}{2}S_{r2}^T\dot H(q)S_{r2}-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+\frac{1}{2}S_{r2}^T(\dot H(q)-2C(q,\dot q))S_{r2}+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^TR_M(\delta\varphi F(t) K_d\bm{sign}(S_{r1}))\delta\varphi F(t) K_d\bm{sign}(S_{r1})+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+v\dot v.\label{eq:V1_1}

@@

Since the command $ -\delta\varphi F(t) K_dS_{r1}$ is bounded, one can select reasonably small positive scalars $c_1$ and $c_2$ such that $0<c_1\le R(\hat\tau_t)<1$ and $0<c_2\le R(\hat\tau_\theta)<1$, and obtain a diagonal matrix as follows:

@@

K_1 = \begin{bmatrix}c_1&0\\0&c_2\end{bmatrix}.\label{eq:saturated ratio}

@@

v\dot v =\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\bm{sign}(S_{r2}),

@@

\dot V&\le -S_{r2}^T\delta\varphi  K_1K_dF(t)\bm{sign}(S_{r2})+S_{r2}^T(\varphi_1+\varphi_2) \bm{sign}(S_{r2})+\varphi\delta S_{r2}^T K_1K_d(F(t)-K_1^{-1})\bm{sign}(S_{r2})\\
&\le \varphi S_{r2}^T\bm{sign}(S_{r2})-\varphi\delta S_{r2}^T K_d\bm{sign}(S_{r2})

@@

owing to the boundedness of the expression above, a large enough $\delta$ can be selected such that $\dot V \le 0$. According to the description of Barbalat¡¯s lemmas, it implies that $S_{r2}$ asymptotically converges to zero with the set of suitable designer parameters.\par
Utilizing $S_{r2}=0$ leads to the equation

@@

and select a Lyapunov function candidate

@@

if and only if $S_{q1} = 0$, accordingly, $S_{q2} = 0$, one has $\dot V_1=0$. Hence the asymptotically stability of Eq.~(\ref{eq:eqSq2}) has been proved, and therefore $S_{q2}\rightarrow 0$ as $t\rightarrow \infty$ produces

@@

Select the Lyapunov function candidate $V_2=\frac{1}{2}\Delta \bm q^T\Delta\bm q$, the derivative of which follows

@@

if and only if $\Delta \bm q = 0$, namely, $\Delta \dot {\bm q} = 0$, one has $\dot V_2=0$. Hence asymptotically convergence of $\Delta \bm q$ has been explained, namely, $\Delta \bm q\rightarrow 0$ as $t\rightarrow \infty$. To conclude, the error state $\Delta \bm q$ of the system~(\ref{eq:dynamics 1}) asymptotically converges to the origin with the limited input~(\ref{eq:slider control}). This completes the proof.

@@

\tau = -\bm{g}(\delta\varphi F(t) K_{d1}\bm{sign}(S_{r3}))

@@

\tau = -\bm{g}(\delta\varphi F(t) K_{d1}\bm{sign}(S_{r4})),

@@

In this section, the simulations of the regulation of the deployment dynamics with saturated sliding mode control, saturated terminal sliding mode control and quasi-saturated terminal sliding mode control, and for clarity, these controllers are abbreviated as S-SMC, ST-SMC and QSF-SMC, respectively. The parameters of these controllers can selected as follows: $\delta=1.1$, $\varphi= 1\times 10^{-3}$, $\alpha = 0.8$, $K_d(K_{*})=1$, $\gamma = 0.1$ and $\hat\alpha=0.6$.\par
Besides the designer parameter determined, the simulation initial and desired condition can be described as follows. The circular orbital altitude is selected around 260 $km$ approximately, and the orbital angular velocity is about $1.17\times 10^{-3}$ $rad/s$. At the initial ejection of the subsatellite, the true anomaly is assumed 0 $rad$. Meanwhile, the initial deployment velocity is about 3 $m/s$, and the subsatellite is going to be deployed at 3.5 $km$ vertically. The pitching and rolling dynamics is varying in a small range, and hence, without loss of generality, the numerical initial condition gives rise to: $\theta_0 = 0.1$, $\varphi_0 = 0.1$, $\lambda_0 = 0.02$; the desired condition is set to be $\theta_d = 0$, $\phi_d = 0$ and $\lambda_d = 1$, accordingly, $\lambda(t_0)\rightarrow\lambda_d$, $t\rightarrow\infty$ with $\theta(t_0)\rightarrow\theta_d$, $\phi(t_0)\rightarrow\phi_d$, $t\rightarrow\infty$. Hence the dimensionless tether length is varying in $[0.02,1]$.\par

@@

At this point, the simulations of the deployment dynamics of the TSS are able to be carried on, and the results are illustrated in Figure~\ref{fig:3} to~\ref{fig:6}. Figure~\ref{fig:3} shows the dynamics of the length in different control laws, and it is obvious that the all the methods finish the deployment in aspect of the tether length. ST-SMC and QST-SMC complete the deployment at about 4 $rad$, and from Figure~\ref{fig:4}, S-SMC and QST-SMC have bigger length rate than ST-SMC before 1.2 $rad$, and hence the curves of the tether length climb more rapidly at the beginning. In contrary, the length rate of ST-SMC doesn't decline dramatically like the other two methods after 1.2 $rad$. This slowly varying length rate yields a smooth trajectory of the deployment shown in Figure~\ref{fig:3}, and meanwhile the length rate converges to origin more early than $S-SMC$ and $QST-SMC$.\par

@@

The inputs in S-SMC for dynamic model~(\ref{eq:dynamic model}) is exhibited in Figure~\ref{fig:6}, in which it is obvious that the limited situations occur frequently. Thanks to the scalable dimensionless transform determined conservatively, the tension exerted on the tether falls in the reasonable region.

@@

>>>>

\bm\tau = -\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r1})),\label{eq:slider control}

@@

\frac{d}{dt}F(t) = \delta\varphi S_{r2}^T K_1e^{-2K_1^{-1}F^{-1}(t)}\bm{sign}(S_{r2})

@@

\textcolor{red}{$\varphi$ is the boundedness of the lumped uncertainty. $\gamma$ is designed as a time-varying positive scalar, $\varphi>1$ is the designer parameter, such that the deployment dynamics can be regulated as desired.}

@@

It is obvious that $S_{q_1}$ is bounded, or exactly saturated, hence $S_{r_1}$ is also bounded if the dynamics can be controlled as desired. \textcolor{red}{In addition, for guaranteeing that $x=\gamma\int^t_{t_0}S_{q1}(\varsigma) d\varsigma$ always falls in the field of definition of $y=arcsin(x)$, $\gamma$ can be designed as a time-varying parameter
    \gamma=\begin{cases}
    \gamma_0&\Vert\int^t_{t_0}S_{q1}(\varsigma) d\varsigma\Vert\le 1\\
    \gamma_0\left[\int^t_{t_0}S_{q1}(\varsigma) d\varsigma\right]^{-1}&\Vert\int^t_{t_0}S_{q1}(\varsigma) d\varsigma\Vert> 1,
    where $\gamma_0$ can be selected as $\gamma_0=0.1$, such that $\gamma\int_{t_0}^tS_{q1}d\varsigma$ is guaranteed to satisfy $\gamma\int_{t_0}^tS_{q1}d\varsigma\in[-1,1]$.}\par

@@

H(q)\dot S_{r2} = -\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r1}))-C(q,\dot q)S_{r2}-Y_{r2}\Theta.

@@

&= -S_{r2}^T\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TC(q,\dot q)S_{r2}+\frac{1}{2}S_{r2}^T\dot H(q)S_{r2}-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+\frac{1}{2}S_{r2}^T(\dot H(q)-2C(q,\dot q))S_{r2}+v\dot v\\
&= -S_{r2}^T\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r1}))+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+v\dot v\\
&= -S_{r2}^TR_M(\delta\varphi F(t) \bm{sign}(S_{r1}))\delta\varphi F(t) \bm{sign}(S_{r1})+S_{r2}^T\bm\tau_d-S_{r2}^TY_{r2}\Theta+v\dot v.\label{eq:V1_1}

@@

Since the command $ -\delta\varphi F(t) S_{r1}$ is bounded, one can select reasonably small positive scalars $c_1$, $c_2$ and $c_3$ such that $0<c_1\le R(\hat\tau_t)<1$, $0<c_2\le R(\hat\tau_\theta)<1$ and $0<c_3\le R(\hat\tau_\phi)<1$, and obtain a diagonal matrix as follows:

@@

K_1 = \begin{bmatrix}c_1&0&0\\0&c_2&0\\0&0&c_3\end{bmatrix}.\label{eq:saturated ratio}

@@

v\dot v =\varphi\delta S_{r2}^T K_1(F(t)-K_1^{-1})\bm{sign}(S_{r2}),

@@

\dot V&\le -S_{r2}^T\delta\varphi  K_1F(t)\bm{sign}(S_{r2})+S_{r2}^T(\varphi_1+\varphi_2) \bm{sign}(S_{r2})+\varphi\delta S_{r2}^T K_1(F(t)-K_1^{-1})\bm{sign}(S_{r2})\\
&\le \varphi S_{r2}^T\bm{sign}(S_{r2})-\varphi\delta S_{r2}^T \bm{sign}(S_{r2})\\
&\le\varphi(1-\delta)S_{r2}^T\bm{sign}(S_{r2})\le 0

@@

where a large enough $\delta>1$ can be selected such that $\dot V \le 0$. According to the description of Barbalat¡¯s lemmas, it implies that $S_{r2}$ asymptotically converges to origin. Also, the finite convergent time of desired sliding surface can be calculated by referring to~\cite{Ma201667}.}\par
\textcolor{red}{Utilizing $S_{r2}=0$ leads to the equation

@@

and for employing Lyapunov second method to analyse the stabilities of $S_{q2}$ and $\Delta\bm q$, we firstly select a Lyapunov function candidate

@@

if and only if $S_{q1} = 0$, accordingly, $S_{q2} = 0$, one has $\dot V_1=0$. Hence the asymptotically stability of Eq.~(\ref{eq:eqSq2}) has been proved, and therefore $S_{q2}\rightarrow 0$ as $t\rightarrow \infty$ induces

@@

We proceed to select the Lyapunov function candidate $V_2=\frac{1}{2}\Delta \bm q^T\Delta\bm q$ for deriving the asymptotically stability by using Lyapunov second method, and the derivative of $V_2$ follows

@@

if and only if $\Delta \bm q = 0$, namely, $\Delta \dot {\bm q} = 0$, one has $\dot V_2=0$. Hence asymptotically convergence of $\Delta \bm q$ has been explained, namely, $\Delta \bm q\rightarrow 0$ as $t\rightarrow \infty$. To conclude, the error state $\Delta \bm q$ of the system~(\ref{eq:dynamics 1}) asymptotically converges to the origin with the limited input~(\ref{eq:slider control})}. This completes the proof.

@@

\tau = -\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r3}))

@@

\tau = -\bm{g}(\delta\varphi F(t) \bm{sign}(S_{r4})),

@@

In this section, the simulations of the regulation of the deployment dynamics with saturated sliding mode control, saturated terminal sliding mode control and quasi-saturated terminal sliding mode control, and for clarity, these controllers are abbreviated as S-SMC, ST-SMC and QSF-SMC, respectively. The parameters of these controllers can selected as follows: $\delta=1.1$, $\varphi= 1\times 10^{-3}$, $\alpha = 0.8$, $\gamma = 0.1$ and $\hat\alpha=0.6$.\par
Besides the designer parameter determined, the simulation initial and desired condition can be described as follows. The circular orbital altitude is selected around 260 $km$ approximately, and the orbital angular velocity is about $1.17\times 10^{-3}$ $rad/s$. At the initial ejection of the subsatellite, the true anomaly is assumed 0 $rad$. Meanwhile, the initial deployment velocity is about 3 $m/s$, and the subsatellite is going to be deployed at 3.5 $km$ vertically. The pitching and rolling dynamics is varying in a small range, and hence, without loss of generality, the numerical initial condition gives rise to: $\theta_0 = 0.1$, $\phi_0 = 0.1$, $\lambda_0 = 0.02$; the desired condition is set to be $\theta_d = 0$, $\phi_d = 0$ and $\lambda_d = 1$, accordingly, $\lambda(t_0)\rightarrow\lambda_d$, $t\rightarrow\infty$ with $\theta(t_0)\rightarrow\theta_d$, $\phi(t_0)\rightarrow\phi_d$, $t\rightarrow\infty$. Hence the dimensionless tether length is varying in $[0.02,1]$.\par

@@

At this point, the simulations of the deployment dynamics of the TSS are able to be carried on. \textcolor{red}{We firstly conduct the simulation about the pure deployment dynamics without any uncertainties to verify the effectiveness of the proposed control laws, and the results are illustrated in Figure~\ref{fig:3} to~\ref{fig:6}.} Figure~\ref{fig:3} shows the dynamics of the length in different control laws, and it is obvious that the all the methods finish the deployment in aspect of the tether length. ST-SMC and QST-SMC complete the deployment at about 4 $rad$, and from Figure~\ref{fig:4}, S-SMC and QST-SMC have bigger length rate than ST-SMC before 1.2 $rad$, and hence the curves of the tether length climb more rapidly at the beginning. In contrary, the length rate of ST-SMC doesn't decline dramatically like the other two methods after 1.2 $rad$. This slowly varying length rate yields a smooth trajectory of the deployment shown in Figure~\ref{fig:3}, and meanwhile the length rate converges to origin more early than S-SMC and QST-SMC.\par

@@

The inputs in S-SMC for dynamic model~(\ref{eq:dynamic model}) is exhibited in Figure~\ref{fig:6}, in which it is obvious that the limited situations occur frequently. Thanks to the scalable dimensionless transform determined conservatively, the tension exerted on the tether falls in the reasonable region.\par

@@

\textcolor{red}{So far, the proposed control laws have been verified in the undisturbed environment. However, space perturbations and parameter uncertainties usually occur during the deployment of the tethered satellite system in engineering application. For checking the robustness of the proposed control laws, based on the dynamics~(\ref{eq:dynamics 1}), we have the following assumptions about uncertainties and disturbances:
\frac{1}{2}&\le\frac{\hat{\bar m}}{\bar m}\le 2\\
\frac{1}{2}&\le\frac{\hat L}{L}\le 2\\
\frac{1}{2}&\le\frac{\hat \Omega}{\Omega}\le 2\\
\bm\tau_d&=\left(0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000),0.05\sin(2\pi t/5000)\right)^T,
where $\hat\bullet$ stands for the measurement of the term $\bullet$ with uncertainties, and $\bm\tau_d$ is the external disturbance meaning  space perturbations, such as gravity gradient torque, aerodynamic interference and solar radiation pressure torque. In this simulation, the measured terms can be specified as: $\hat{\bar m}=1.1\bar m$ may mean some space debris affixed to the subsatellite; $\bar L=1.02L$ and $\bar \Omega=0.98\Omega$ mean some measurement errors happening to the total length and the orbital angular velocity. In Fig~\ref{fig:6}, although the dynamics has been already stable, the tension is switching in the high frequency. It is commonplace in sliding mode controller design, called chattering, since the signum function is used in the input, which may cause unmodelled high frequency dynamics, and for removing chattering, the boundary layer technique is introduced into the proposed control laws through replacing $sign(x)$ by $sat(x/\nu)$, where
sat(x/\nu)=
x/\nu,&\vert x\vert <\nu\\
sign(x),&\vert x\vert\ge\nu
and $\nu$ is the thickness of the boundary layer specified as a positive scalar $\nu=1\times 10^{-4}$ in this simulation.}\par
The simulation results of the deployment dynamics subjected to uncertainties are shown from Fig~\ref{fig:7} to Fig~\ref{fig:10}. In Fig~\ref{fig:7}, although there exist the parameter uncertainties, QST-SMC still performs the best deployment among the three methods, and the tethers are deployed to the desired position at about 8 $rad$ by observing the dynamics of the in-plane and out-of-plane angles from Fig~\ref{fig:9}. In addition, the result in Fig~\ref{fig:10} shows that the tension falls in the reasonable region, and meanwhile the chattering of the input in the stable status is eliminated by utilizing the boundary layer technique. Hence, the robustness of the proposed control laws can guarantee the stability of the deployment subjected to uncertainties and disturbances.}
\includegraphics[width=0.6\textwidth]{paper4_fig7_20161025.eps}
\caption{Length dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:7}
\includegraphics[width=0.6\textwidth]{paper4_fig8_20161025.eps}
\caption{Length rate dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:8}
\includegraphics[width=0.6\textwidth]{paper4_fig9_20161025.eps}
\caption{Angle dynamics of the deployment subjected to uncertainties and disturbances}
\label{fig:9}
\includegraphics[width=0.6\textwidth]{paper4_fig10_20161025.eps}
\caption{Tension produced by utilizing boundary layer technique}
\label{fig:10}

### 2016-10-26 paper4.tex
20161026

% Some useful packages...
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
%\usepackage[figurename=Fig.,labelfont=bf,labelsep=period]{caption}
\usepackage{subcaption}
%\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
%\usepackage{newtxtext,newtxmath}
\usepackage[colorlinks=true,citecolor=red,linkcolor=black]{hyperref}
%

@@

\author[2]{Zhiqiang Ma}
\author[1]{Guanghui Sun}
\author[2]{Zhengkai Li}

\affil[1]{Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China, with corresponding author email. Email: guanghuisun@hit.edu.cn}
\affil[2]{Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China, with corresponding author email.}
%\affil[2]{Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China}
%\affil[2]{Third affiliation address}

>>>>

% Some useful packages...
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{graphicx}
%\usepackage[figurename=Fig.,labelfont=bf,labelsep=period]{caption}
\usepackage{subcaption}
%\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsbsy}
%\usepackage{newtxtext,newtxmath}
\usepackage[colorlinks=true,citecolor=black,linkcolor=black]{hyperref}
%

@@

\author[1]{Zhiqiang Ma}
\author[2]{Guanghui Sun}
\author[3]{Zhengkai Li}
\affil[1]{Ph.D. Candidate, Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China, Email: zhiqiangma@hit.edu.cn}
\affil[2]{Associate Professor, Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China, Email: guanghuisun@hit.edu.cn}

\affil[3]{Ph.D. Candidate, Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China, Email: lizhengkai1@126.com}
%\affil[2]{Research Institute of Intelligent Control and Systems, Harbin Institute of Technology, Harbin 150001, China}
%\affil[2]{Third affiliation address}
